|

### 2021-08-10 prova_knit_finale.Rmd
arrow

\subsubsection{Convergence of the IS estimator}\label{pf_is_conv}
We can show that the auto-normalized IS estimator is both consistent and asymptotically normal.\
Consider the auto-normalized IS estimator $\hat \varphi= \sum_{n=1}^N\varphi (X^n)W^n$. We can write:
\begin{equation*}
   \hat \varphi_{AN} - \E_{\Q}(\varphi)=\E_{\Q^N}(\varphi)-\E_{\Q}(\varphi)=\frac{N^{-1}\sum_{n=1}^Nw(X^n)[\varphi(X^n)-\E_{\Q}(\varphi)]}{N^{-1}\sum_{n=1}^Nw(X^n)}
\end{equation*}
Let $\bar \varphi(X):=\varphi(X)-\E_{\Q}(\varphi)$. Applying the CLT to the numerator, assuming that $\E_{\M}(w^2\bar\varphi^2)<+\infty$, we get:
\begin{equation}
    \sqrt{N}\Bigg\{\frac{1}{N}\sum_{n=1}^Nw(X^n)\bar \varphi (X^n)\Bigg\}\implies \N \big(0,\E_{\M}(w^2\bar\varphi^2)\big)\label{is_conv_clt}
\end{equation}
where $\implies$ denotes convergence in distribution. Instead, for the denominator the strong LLN implies that:
\begin{equation}
    \frac{1}{N}\sum_{n=1}^Nw(X^n)\rightarrow^{\text{a.s.}} \E_{\M}(w) \label{is_conv_lln}
\end{equation}
Applying Slutsky's theorem, together with \eqref{is_conv_clt} and \eqref{is_conv_lln}, we obtain:
\begin{equation}
   \sqrt{N}\big[  \hat \varphi_{AN} - \E_{\Q}(\varphi)\big] \implies \N \Bigg(0,\frac{\E_{\M}(w^2\bar\varphi^2)}{[\E_{\M}(w)]^2}\Bigg)
\end{equation}
Therefore, the IS (auto-normalized) estimator for $\varphi$ is asymptotically normal. Consistency can be proved in an analogous way, applying the strong LLN and lemma \ref{lemma_imp_sam_1}:
\begin{equation}
    \frac{1}{N}\sum_{n=1}^Nw(X^n) \varphi (X^n)\rigtharrow^{\text{a.s.}}  \E_\M(\varphi \cdot w)=\E_\Q(\varphi)\cdot \E_\M(w)
\end{equation}
Together with \eqref{is_conv_lln}, we get strong convergence of the estimator (that implies consistency).
\begin{equation}
   \hat \varphi_{AN}=\sum_{n=1}^NW^n \varphi (X^n)\rigtharrow^{\text{a.s.}}  \E_\Q(\varphi)
\end{equation}

@@

\subsection{Implementation}
Consider again the random walk plus noise model of section \ref{implementation},the challenge here is to estimate the filtered states using a Sequential Importance Sampling algorithm. The main idea of the SIS applied to this univariate linear gaussian model is described in the following algorithm. We indicate with $n$ the sample size of time observations and with $N$ the generated sample size for each step of the Sequenial Monte Carlo. 
\begin{algorithm} SIS filter for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarize $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(x_{t-1}^{(i)},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t})$
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
Our \code{SISfun} function implemented in R replicate this steps.
\
\hrule
\hrule
\
\code{SISfun(data,N,m0,C0,tau,sigma)}
\
\hrule
\textbf{Arguments}

>>>>

\subsubsection{Convergence of the IS estimator}\label{pf_is_conv}
We can show that the auto-normalized IS estimator is both consistent and asymptotically normal.\
Consider the auto-normalized IS estimator $\hat \varphi= \sum_{n=1}^N\varphi (X^n)W^n$. We can write:
\begin{equation*}
   \hat \varphi_{AN} - \E_{\Q}(\varphi)=\E_{\Q^N}(\varphi)-\E_{\Q}(\varphi)=\frac{N^{-1}\sum_{n=1}^Nw(X^n)[\varphi(X^n)-\E_{\Q}(\varphi)]}{N^{-1}\sum_{n=1}^Nw(X^n)}
\end{equation*}
Let $\bar \varphi(X):=\varphi(X)-\E_{\Q}(\varphi)$. Applying the CLT to the numerator, assuming that $\E_{\M}(w^2\bar\varphi^2)<+\infty$, we get:
\begin{equation}
    \sqrt{N}\Bigg\{\frac{1}{N}\sum_{n=1}^Nw(X^n)\bar \varphi (X^n)\Bigg\}\implies \N \big(0,\E_{\M}(w^2\bar\varphi^2)\big)\label{is_conv_clt}
\end{equation}
where $\implies$ denotes convergence in distribution. Instead, for the denominator the strong LLN implies that:
\begin{equation}
    \frac{1}{N}\sum_{n=1}^Nw(X^n)\rightarrow^{\text{a.s.}} \E_{\M}(w) \label{is_conv_lln}
\end{equation}
Applying Slutsky's theorem, together with \eqref{is_conv_clt} and \eqref{is_conv_lln}, we obtain:
\begin{equation}
   \sqrt{N}\big[  \hat \varphi_{AN} - \E_{\Q}(\varphi)\big] \implies \N \Bigg(0,\frac{\E_{\M}(w^2\bar\varphi^2)}{[\E_{\M}(w)]^2}\Bigg)
\end{equation}
Therefore, the IS (auto-normalized) estimator for $\varphi$ is asymptotically normal. Consistency can be proved in an analogous way, applying the strong LLN and lemma \ref{lemma_imp_sam_1}:
\begin{equation}
    \frac{1}{N}\sum_{n=1}^Nw(X^n) \varphi (X^n)\xrigtharrow{\text{a.s.}}  \E_\M(\varphi \cdot w)=\E_\Q(\varphi)\cdot \E_\M(w)
\end{equation}
Together with \eqref{is_conv_lln}, we get strong convergence of the estimator (that implies consistency).
\begin{equation}
   \hat \varphi_{AN}=\sum_{n=1}^NW^n \varphi (X^n)\xrigtharrow{\text{a.s.}}  \E_\Q(\varphi)
\end{equation}

@@

\subsection{Implementation}
Consider again the random walk plus noise model of section \ref{implementation}, the challenge here is to estimate the filtered states using a Sequential Importance Sampling algorithm. The main idea of the SIS applied to this univariate linear gaussian model is described in the following algorithm. We indicate with $n$ the sample size of time observations and with $N$ the generated sample size for each step of the Sequenial Monte Carlo. 
\begin{algorithm} SIS filter for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarize $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(x_{t-1}^{(i)},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t})$
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
Our \code{SISfun} function implemented in R replicate this steps.
\
\hrule
\hrule
\
\code{SISfun(data,N,m0,C0,tau,sigma)}
\
\hrule
\textbf{Arguments}
